My title

### 2024-11-26 paper/paper.qmd
update paper result

Some scholars have examined the changing climate. @xu_melting_2009 analyze the effects of rising temperatures in the Himalayas, highlighting increased frequency and duration of extreme events and shifts in ecosystems. These changes pose challenges to water supply, agriculture, and human populations. @visser_eliminating_2021 investigates the relationship between precipitation and temperature using data from the Australian Bureau of Meteorology. Visser's regression model indicates that average precipitation intensities increase with temperature, suggesting more intense rainfall in a warmer climate. The role of sea level pressure is also significant. @wills_systematic_2022 note that observed trends in sea level pressure have intensified warming in the Indo-Pacific Warm Pool and caused slight cooling in the eastern equatorial Pacific. However, as @zhang_economic_2017 argue, much of the research has focused on temperature and precipitation. @zhang_economic_2017 expands on this by incorporating additional predictors—relative humidity and wind speed—and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.

Temperatures significantly impact airport operations. Rising temperatures significantly affect aircraft performance, potentially leading to take-off weight restrictions and the need for longer runways. This directly impacts airport capacity and operations [@coffel_climate_2015]. Temperature forecast models vary in different locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].

@@

## MLR Model

@@

The purpose of including polynomial terms in this model is to capture non-linear relationships between the predictors and the response variable, allowing the model to fit more complex, curved patterns that a purely linear model cannot accommodate. By incorporating quadratic terms for each predictor, the model gains flexibility in representing real-world data relationships, reducing bias, and addressing systematic patterns in residuals that may indicate non-linearity. This enhances the model's predictive performance, making it capable of explaining variance that would otherwise remain unaccounted for in a simple linear regression, while still maintaining the interpretability and simplicity of a linear regression framework.

@@

The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training. The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.

@@

The polynomial linear regression model was selected over alternative non-linear methods due to its balance between complexity and simplicity. Polynomial terms extend the linear model's capacity to capture non-linear patterns while preserving the interpretability of regression coefficients. The inclusion of transformations, such as logarithmic and Box-Cox, further improves the model by addressing issues of skewness and heteroscedasticity observed in the raw data.

Finally, the model assumptions were rigorously evaluated in @sec-validation. Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals. Variance Inflation Factor (VIF) values for the predictors were all well below the critical threshold, indicating no significant multicollinearity. These results ensure that the model remains valid and reliable for inference and prediction.

@@

## Model Metrics: Explains 61% Variance with Low RMSE
The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**. These values demonstrate that approximately 61% of the variability in $\log(\text{mean\_temp})$ is explained by the predictors, even after accounting for the complexity of the model.

@@

@@

## The finding does not agree with Visser
## the finding does match with Wills

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

@@

While the polynomial linear regression model captures much of the relationship between the predictors and the log-transformed mean temperature, some limitations remain. @fig-assumption-1 indicates that the relationship is not completely linear, with minor deviations suggesting that the model may not fully account for certain non-linearities. Additionally, @fig-assumption-3 shows slight evidence of heteroscedasticity at the extreme fitted values, indicating that the variance of residuals is not entirely constant. These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity. 
Weaknesses and next steps should also be included.

@@

>>>>

Some scholars have examined the changing climate. @xu_melting_2009 analyze the effects of rising temperatures in the Himalayas, highlighting increased frequency and duration of extreme events and shifts in ecosystems. These changes pose challenges to water supply, agriculture, and human populations. @visser_eliminating_2021 investigates the relationship between precipitation and temperature using data from the Australian Bureau of Meteorology. Visser's regression model indicates that average precipitation intensities increase with temperature, suggesting more intense rainfall in a warmer climate. The role of sea level pressure is also significant. @wills_systematic_2022 note that observed trends in sea level pressure have caused slight cooling in the eastern equatorial Pacific. However, as @zhang_economic_2017 argue, much of the research has focused on temperature and precipitation. @zhang_economic_2017 expands on this by incorporating additional predictors—relative humidity and wind speed—and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.

Temperatures significantly impact airport operations. Rising temperatures significantly affect aircraft performance, leading to take-off weight restrictions and the need for longer runways. This directly impacts airport capacity and operations [@coffel_climate_2015]. Temperature forecast models vary in different locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].

@@

## MLR Polynomial Model

@@

The purpose of including polynomial terms in this model is to capture non-linear relationships between the predictors and the response variable, allowing the model to fit more complex, curved patterns that a purely linear model cannot accommodate. This enhances the model's predictive performance, making it capable of explaining variance that would otherwise remain unaccounted for in a simple linear regression, while still maintaining the interpretability and simplicity of a linear regression framework.

@@

The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2. The $R^2$ value is calculated using the following formula:

\begin{align}
R^2 &= 1 - \frac{\text{SSR}}{\text{TSS}} \\
\text{SSR} &= \sum_{i=1}^{n} (y_i - \hat{y}_i)^2 \\
\text{TSS} &= \sum_{i=1}^{n} (y_i - \bar{y})^2
\end{align}

- $\text{SSR}$ (Sum of Squared Residuals) measures the total squared difference between the observed values ($y_i$) and the predicted values ($\hat{y}_i$).
- $\text{TSS}$ (Total Sum of Squares) measures the total squared difference between the observed values ($y_i$) and their mean ($\bar{y}$).
- $R^2$ quantifies the proportion of the variance in the response variable explained by the model.

 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training. The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.

@@

The polynomial linear regression model was selected over alternative non-linear methods due to its balance between complexity and simplicity. Polynomial terms extend the linear model's capacity to capture non-linear patterns while preserving the interpretability of regression coefficients. The inclusion of transformations, such as logarithmic and Box-Cox, further improves the model by addressing issues of skewness and heteroscedasticity observed in the raw data.  By incorporating quadratic terms for each predictor, the model gains flexibility in representing real-world data relationships, reducing bias, and addressing systematic patterns in residuals that may indicate non-linearity.

Finally, the model assumptions were evaluated in @sec-validation. Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals. Variance Inflation Factor (VIF) values for the predictors were all well below the critical threshold, indicating no significant multicollinearity. These results ensure that the model remains valid and reliable for inference and prediction.

@@

## Model Metrics: Explains 61% Variance with Low RMSE {#sec-result-r2}
The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**. These values demonstrate that approximately 61% of the variability in $\log(\text{mean\_temp})$ is explained by the predictors, even after accounting for the complexity of the model.

@@

## Predicted vs. Actual Plot shows good overall fit

```{r}
#| label: fig-predictedvsfit
#| fig-cap: Predicted vs. Actual plot shows good fit
#| fig-subcap: ['The model is accurate in the train data', 'The model is accurate in the test data']
#| echo: false
#| warning: false
#| layout-ncol: 2
#| 
train_data$predicted <- predict(glm, newdata = train_data)
test_data$predicted <- predict(glm, newdata = test_data)

ggplot(train_data, aes(x = log_mean_temp, y = predicted)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted vs Actual Log Mean Temperature",
       x = "Actual Log Mean Temperature",
       y = "Predicted Log Mean Temperature") +
  theme_minimal()
ggplot(test_data, aes(x = log_mean_temp, y = predicted)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted vs Actual Log Mean Temperature",
       x = "Actual Log Mean Temperature",
       y = "Predicted Log Mean Temperature") +
  theme_minimal()
```

As displayed in @fig-predictedvsfit, points aligning closely along the diagonal line indicate that the predicted log-transformed mean temperature matches well with the observed values. In @fig-predictedvsfit-1, The points closely align along the red diagonal line, indicating that the model predictions are highly accurate for the training dataset. This alignment suggests that the model captures the relationship between the predictors and the log-transformed mean temperature effectively, with minimal bias or variance issues within the training data. In @fig-predictedvsfit-2, the points on the test data plot also align reasonably well along the red diagonal line, though with slightly more scatter compared to the training data. This result indicates that the model generalizes well to unseen data, maintaining good predictive accuracy and avoiding overfitting. The consistency between the train and test plots suggests the model has a good balance between complexity and generalization. In summary, the Predicted vs. Actual plots provide clear visual evidence of the model's accuracy and robustness for both training and test datasets.

@@

## The finding does not agree with Visser

## the finding does match with Wills

@@

While the polynomial linear regression model captures much of the relationship between the predictors and the log-transformed mean temperature, some limitations remain. @fig-assumption-1 indicates that the relationship is not completely linear, with minor deviations suggesting that the model may not fully account for certain non-linearities. Additionally, @fig-assumption-3 shows slight evidence of heteroscedasticity at the extreme fitted values, indicating that the variance of residuals is not entirely constant. These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity.

@@

### Limitation on ECCC's methodology
